\documentclass{article}
\usepackage{parskip}

\title{Phase 1 - September: Advantages of parallel compared with serial computation, for route-planning problems and how increased data size may affect this.}

\begin{document}
\maketitle

\subsection*{ Phase 1 Project Selection Status Report}

Name: Marcus Alexander Karmi September

College: Clare

CRSID: maks2

Director of Studies: Lawrence Paulson

\subsection*{Please complete 1, 2 and 3 below.}

\begin{enumerate}
    \item \textit{Please write 100 words on your current project ideas.}

         It's possible to solve APSP through repeated squaring of
         the adjacency matrix. Since parallelising matrix multiplication is a
         well-explored area, APSP may be solved more efficiently on a multi-core system
         by using efficient parallel matrix multiplication techniques, compared to
         serial alternatives.

         I'll implement such a parallel algorithm in Java, using efficient
         matrix multiplication techniques that minimise data movement, as well
         as  other optimizations. With Java's Thread API, I can simulate a
         processor with an arbitrary amount of cores, interconnected through
         arbitrary network topologies. I'll evaluate the advantage of my
         parallel implementation, and then see how this advantage and the
         design of the algorithm change as the data size increases relative to
         the number of processing elements available.

     \item \textit{Please list names of potential project supervisors, indicating
             any interactions you have had with them, for example: not
         contacted, awaiting reply, in discussion, agreed to supervise.}

         Dr Jagdish J. Modi (jm505):
         \begin{itemize}
             \item Agreed to supervise
             \item I have been in contact with Dr J. Modi through several email exchanges
                 and we have had two Zoom calls.
         \end{itemize}


\item \textit{Is there any chance that your project will involve any
computing resources other than the Computing Service's MCS and
software that is already installed there, for example: your own
machine, machines in College, special peripherals, imported
software packages, special hardware, network access, substantial
extra disc space on the MCS. If so indicate below what, and what it is needed for.}

\begin{itemize}
    \item My personal laptop, needed for running my Java programs
    \item Real-world graph datasets, needed for evaluating my algorithm
\end{itemize}
\end{enumerate}

\end{document}
