%! TEX root = diss.tex
\documentclass[../diss.tex]{subfiles}
\chapter{Conclusions}

In this dissertation, I have described the design, implementation and
evaluation of a parallel system simulator and an algorithm for solving
\ac{APSP}.  The project was a great success: The simulator was simple,
expressive and efficient and the \ac{MatSquare} algorithm and its extensions
gave correct outputs for large graphs. Furthermore, the evaluation covered
an elaborate qualitative analysis on the benefit of parallel computation, which
contributed to the original aims of the project---investigating the benefit of
parallel computation and its extent on different parallel system and inputs.

% Results
% Results {{{
\section{Results}%
\label{sec:Results}

As shown in \autoref{sec:Parallel simulation}, the project has met all the
success criteria related to developing an efficient simulation of a parallel
system. A parallel algorithm, \ac{MatSquare}, was developed using the interface
the simulator provided. This algorithm met all the success criteria that was
initially outlined, and exceeded these through generalisation and
further optimisations.
Additionally, the analysis of the algorithm's execution proved the final
success criteria, and went beyond by considering different classes of
parallel computers.

% }}}

% Lessons learnt {{{
\section{Lessons learnt}%
\label{sec:Lessons learnt}

Using the proper tools, such as an IDE for writing Java code and \texttt{git} for
doing revision control, helped immensely with keeping implementation productive.
Additionally, creating UML diagrams before writing any code also made managing
the large project easier, allowing more time for writing code rather than trying to
figure out how components piece together.

During development, I was often fixated on minor aspects of the parallel
simulation interface, which turned out less necessary than expected.
Otherwise, the whole approach was educational and I learned a lot about planning
a large project and writing parallel programs.

% }}}

% Future work {{{
\section{Future work}%
\label{sec:Future work}

The input graphs used in the dissertation were very sparse. This could be exploited
to reduce the time complexity by using parallel matrix multiplication techniques
designed for sparse matrices, such as what \citeauthor{sparseMatMul} propose
\cite{sparseMatMul}.

I also only considered one algorithm for solving \ac{APSP}. It would be interesting
to implement others on the simulator 
and compare how they scale on different types of parallel systems.

Another interesting extension would be to implement the algorithm on real
parallel hardware, such as a \ac{FPGA}.

Furthermore, the algorithm could be applied to Internet routing, where there are
different metrics such as bandwidth and latency to optimize.

Lastly, it would be interesting to explore the benefit of the graph compression and
decompression algorithm for solving other graph optimization problems, such as the
minimum spanning tree problem and the maximum flow problem.

% }}}




